---
title: "Replicate JQT 2011 Xie"
author: "Tahir and Inez"
date: "6 December 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-packages, include=FALSE, eval=FALSE}
  install.packages("tidyverse") #for tibbles.
  install.packages("purrr") 
  install.packages("dbplyr") 
  install.packages("ggplot2")
  install.packages("tidyr")
```

```{r load-packages, include=TRUE, eval=TRUE, message=FALSE}
  require("tidyverse") #for tibbles.
  require("purrr") 
  require("dbplyr") 
  require("ggplot2")
  require("tidyr")
```

# Intoduction
In still code we reproduce the results from "Two MEWMA Charts for Gumbel's Biviate Exponantial Distribution" by Xie et al. (2011) in JQT.

# Data model: GBE data - bivariate
Here we code the GBE data model denoted by $GBE(\theta_1,\theta_2,\delta)$. First we set the parameters and next provide a function to  simulate data from it.
```{r data-model-GBE}
# general random data from bivariate GBE model  
r.gbe <- function(n, t1, t2, d){
  
  U = runif(n,min=0,max=1)
  M = rbinom(n,size=1,prob=d)
  V1 = rexp(n, rate=1)
  V2 = rexp(n, rate=1)
  V = V1 + M*V2
  X1 = t1*(U^d)*V
  X2 = t2*((1-U)^d)*V 
  X = matrix(data = NA,nrow=2,ncol=n)
  X[1,] = X1
  X[2,] = X2
  return(X)
}
```



# Set up monitoring methods
Here we set up the five monitoring methods
```{r functions-monitoring-methods}

#mewma.raw
mewma.raw <- function(sim,r,h,mu.x,sigma.x.inv,dTheta1,dTheta2,dDelta){
   vRL = rep(0,sim)
   vTS = rep(0,sim)
   for (i in 1:sim){
    vZ_prev <- 0
    dE2 <- 0
    while(dE2 < h ){
      vX <- r.gbe(1,dTheta1,dTheta2,dDelta) #generate data
      vZ = r*(vX-mu.x)+(1-r)*vZ_prev
      vZ_prev = vZ
      dE2 = ((2-r)/r) * t(vZ) %*% sigma.x.inv %*% vZ
      vRL[i] = vRL[i] + 1
      vTS[i] = vTS[i] + max(vX)
    }
   }
   ARL = mean(vRL)
   ATS = mean(vTS)
   return(c(ARL,ATS))
}


#ewma.raw
ewma.raw <- function(sim,r,L,mu.x,sigma.x,dTheta1,dTheta2,dDelta){
   vRL = rep(0,sim)
   vTS = rep(0,sim)
   vUCL =  mu.x+L * diag(sigma.x) * sqrt(r/(2-r))
   for (i in 1:sim){
      vZ_prev <- c(0,0)
      vZ <- c(0,0)
      dummy.rl = rep(0,2)
      dummy.ts =rep(0,2)
      
      while(all((abs(vZ) < vUCL))){ #only contiunue if all univariate < ucl
        vX <- r.gbe(1,dTheta1,dTheta2,dDelta) #generate data
        vZ = r*vX+(1-r)*vZ_prev
        vZ_prev = vZ
        vRL[i] = vRL[i] + 1
        dummy.ts[1] = dummy.ts[2] + vX[1]
        dummy.ts[2] = dummy.ts[2] + vX[2]
      } 
    vTS[i] = mean(dummy.ts)
    }
   ARL = mean(vRL)
   ATS = mean(vTS)
   return(c(ARL,ATS))
}


```



Next we run the code for R simulation runs
```{r run-monitoring-methods}
#initialize GBE model
dTheta1 = 1
dTheta2 = 1
dDelta = 0.5
mu.x <- matrix(1,nrow=2,ncol=1)
sigma.x <- matrix(NA,nrow=2,ncol=2)
sigma.x[1,1] = dTheta1^2
sigma.x[1,2] = sigma.x[2,1] = dTheta1*dTheta2*2*gamma(dDelta+1)/gamma(2*dDelta+1)-dTheta1*dTheta2
sigma.x[2,2] = dTheta2^2
sigma.x.inv = solve(sigma.x)
results = data.frame(mewma.raw=double(length=2),
                 ewma.raw = double(length=2),
                 mewma.tra =double(length=2),
                 emwa.tra = double(length=2),
                 tt = double(length=2))
row.names(results) <- c("ARL","ATS")

#set initial values
r <- 0.1
h <- 10.35
L <- 2.901
R=1000

#>>>>>  run simulation   <<<<<<<
#in-control results
results$mewma.raw[1:2] <- mewma.raw(sim=R,r,h,mu.x,sigma.x.inv,dTheta1,dTheta2,dDelta)
                                      
results$ewma.raw[1:2] <- ewma.raw(sim=R,r,L,mu.x,sigma.x,dTheta1,dTheta2,dDelta)

#display results
round(results, digits=0)
```











